Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-15 2017-Xudong-SMT/FM-2018-extended.tex
dessin et listing

\usepackage[T1]{fontenc}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}

@@

\begin{abstract}
Open pNets are used to model the behavior of open systems, both
synchronous or asynchronous, expressed in various calculi or
languages. They are endowed with a symbolic operational semantics in
terms of so-called ``Open Automata''. This allows to check properties of
such systems in a compositionnal manner. 
We implement an algorithm computing this semantics,
building predicates expressing the synchronization conditions
between the events of the pNet sub-systems. Checking such
predicates requires symbolic reasoning over first order logics, but
also over application-specific data. We use the Z3 SMT engine to check
satisfiability of the predicates, and prune the open automaton of its
unsatisfiable transitions.
As an industrial oriented use-case, we use so-called "architectures"
of BIP systems, that have been used to specify the control software of
ESA microsatellites. We use pNets to encode BIP architectures,
extended with explicite data in synchronisations, and we compute their
open automaton semantics. Cet automate peut alors être utilisé pour
monter les propriétés comme, \TODO{ par exemple ..}

@@

 \begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/BIPspec-ArchFailureTimerMax}}
   \caption{FailureTimer Architecture in BIP graphical syntax}  \label{schema:BIParchitecture}
 \end{figure}

\begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/BIPpnet-ArchFailureTimerMax}}
   \caption{FailureTimer pNet encoding}  \label{schema:BIParchitecture}
 \end{figure}

@@

\section{Full examples}
\TODO{May be we use two architecture examples here, a simpler one,
  before the Failure Timer ?}
\subsection{A very simple architecture}
\subsection{Full input for the FailureTimer architecture}
\subsection{Full output for the FailureTimer open automaton}
\subsection{Scaling up: several subsystems on the same bus}

>>>>

\usepackage[T1]{fontenc}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{a4paper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}

@@

\begin{abstract}
Open pNets are used to model the behavior of open systems, both
synchronous or asynchronous, expressed in various calculi or
languages. They are endowed with a symbolic operational semantics in
terms of so-called ``Open Automata''. This allows to check properties of
such systems in a compositionnal manner. 
We implement an algorithm computing this semantics,
building predicates expressing the synchronization conditions
between the events of the pNet sub-systems. Checking such
predicates requires symbolic reasoning over first order logics, but
also over application-specific data. We use the Z3 SMT engine to check
satisfiability of the predicates, and prune the open automaton of its
unsatisfiable transitions.
As an industrial oriented use-case, we use so-called "architectures"
of BIP systems, that have been used to specify the control software of
ESA microsatellites. We use pNets to encode BIP architectures,
extended with explicite data in synchronisations, and we compute their
open automaton semantics. This automaton may then be used to prove
bahavioral properties, like e.g., \TODO{ par exemple ..}

@@

 %% \begin{figure}[t]
 %%   \centerline{\includegraphics[width=12cm]{XFIG/BIPspec-ArchFailureTimerMax}}
 %%   \caption{FailureTimer Architecture in BIP graphical syntax}  \label{schema:BIParchitecture}
 %% \end{figure}

\begin{figure}[t]
   \centerline{\includegraphics[width=12cm]{XFIG/FailureTimerPNET}}
   \caption{FailureTimer pNet encoding}  \label{schema:BIParchitecture}
 \end{figure}

@@

\section{Full examples}
\TODO{May be we use two architecture examples here, a simpler one,
  before the Failure Timer ?}
\subsection{A very simple architecture}
\subsection{Full input for the FailureTimer architecture}
\subsection{Full output for the FailureTimer open automaton}

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, frame=single]
    {s0---failure(true)-->s1, t0---start(true)-->t1}, {---:hb_B:13:1-->},
    (failure(true)=failure(b1:sva_SV0:1:1))/\(start(true)=start(b2:sva_SV0:1:1))
    /\(:hb_B:13:1=fail(b0:sva_SV0:1:1))/\((b1:sva_SV0:1:1=b2:sva_SV0:1:1)
    /\(~(b1:sva_SV0:1:1\/b2:sva_SV0:1:1)\/b0:sva_SV0:1:1))/\(_fail_=:ra:1:1),   {t := 4}
OT ------------------------------------------------------------
        <s0_t0>-----:ra:1:1----><s1_t1>

        {s0---resume(false)-->s0, t0---resume(false)-->t0},   {---:hb_B:16:2-->},
        (resume(false)=resume(b1:sva_SV1:1:2))/\(resume(false)=resume(b2:sva_SV1:1:2))
        /\(:hb_B:16:2=resume(b0:sva_SV1:1:2))/\((b1:sva_SV1:1:2=b2:sva_SV1:1:2)
        /\(~(b1:sva_SV1:1:2\/b2:sva_SV1:1:2)\/b0:sva_SV1:1:2))/\(_resume_=:ra:1:2),   {}
OT ------------------------------------------------------------
        <s0_t0>-----:ra:1:2----><s0_t0>

    {},   {---:hb_B:118:6-->},   (:hb_B:118:6=finish)/\(finish=:ra:1:6),   {}
OT ------------------------------------------------------------
        <s0_t0>-----:ra:1:6----><s0_t0>

        {s1---failure(false)-->s1, t1---start(false)-->t1},   {---:hb_B:13:1-->},
        (failure(false)=failure(b1:sva_SV0:1:1))/\(start(false)=start(b2:sva_SV0:1:1))
        /\(:hb_B:13:1=fail(b0:sva_SV0:1:1))/\((b1:sva_SV0:1:1=b2:sva_SV0:1:1)
        /\(~(b1:sva_SV0:1:1\/b2:sva_SV0:1:1)\/b0:sva_SV0:1:1))/\(_fail_=:ra:1:1),   {}
OT ------------------------------------------------------------
        <s1_t1>-----:ra:1:1----><s1_t1>

        {s1---resume(true)-->s0, t1---resume(true)-->t0},   {---:hb_B:16:2-->},
        (resume(true)=resume(b1:sva_SV1:1:2))/\(resume(true)=resume(b2:sva_SV1:1:2))
        /\(:hb_B:16:2=resume(b0:sva_SV1:1:2))/\((b1:sva_SV1:1:2=b2:sva_SV1:1:2)
        /\(~(b1:sva_SV1:1:2\/b2:sva_SV1:1:2)\/b0:sva_SV1:1:2))/\(_resume_=:ra:1:2),   {}
    OT ------------------------------------------------------------
        <s1_t1>-----:ra:1:2----><s0_t0>

        {t1---tick-->t1},   {},   (t!=0)/\(_tick_=:ra:1:3),   {t := (t-1)}
    OT ------------------------------------------------------------
        <s1_t1>-----:ra:1:3----><s1_t1>

        {},   {---:hb_B:118:6-->},   (:hb_B:118:6=finish)/\(finish=:ra:1:6),   {}
    OT ------------------------------------------------------------
        <s1_t1>-----:ra:1:6----><s1_t1>

        {s1---timeout-->s2, t1---timeout-->t0},   {},   (t=0)/\(ask=:ra:1:4),   {}
    OT ------------------------------------------------------------
        <s1_t1>-----:ra:1:4----><s2_t0>

        {s2---resume(false)-->s2, t0---resume(false)-->t0},   {---:hb_B:16:2-->},
        (resume(false)=resume(b1:sva_SV1:1:2))/\(resume(false)=resume(b2:sva_SV1:1:2))
        /\(:hb_B:16:2=resume(b0:sva_SV1:1:2))/\((b1:sva_SV1:1:2=b2:sva_SV1:1:2)
        /\(~(b1:sva_SV1:1:2\/b2:sva_SV1:1:2)\/b0:sva_SV1:1:2))/\(_resume_=:ra:1:2),   {}
    OT ------------------------------------------------------------
        <s2_t0>-----:ra:1:2----><s2_t0>

        {s2---reset-->s0},   {},   (reset=:ra:1:5),   {}
    OT ----------------------------------------------------------
        <s2_t0>-----:ra:1:5----><s0_t0>

        {},   {---:hb_B:118:6-->},   (:hb_B:118:6=finish)/\(finish=:ra:1:6),   {}
    OT ------------------------------------------------------------
        <s2_t0>-----:ra:1:6----><s2_t0>
\end{lstlisting}

\subsection{Scaling up: several subsystems on the same bus}
